A Data-Driven Framework for Neural Field Modelling

### 2010-11-18 ltx/first_order_framework/LaTeX_Manuscript.tex
deleted all the figures
changed figure to Fig

\subsection{Integro-Difference Equation Neural Field Model}
The combination of modelling techniques in this paper leads to a large amount of notation, so a reference of the symbols used is provided in Table~\ref{tab:Notation}. The model relates the average number of action potentials $g(\mathbf{r},t)$ arriving at time $t$ and position $\mathbf{r}$ to the local post-synaptic membrane voltage $v(\mathbf{r},t)$. The post-synaptic potentials generated at a neuronal population at location $\mathbf{r}$ by action potentials arriving from all other connected populations is described by 
\begin{equation}
	\label{SpikesToPotential} v\left( {\mathbf{r},t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right) \, dt'}. 
\end{equation}
The post-synaptic response kernel $h(t)$ is described by 
\begin{equation}
	\label{SynapticRespKernel} h(t) = \eta(t)\exp{\left(-\zeta t\right)}, 
\end{equation}
where $\zeta=\tau^{-1}$, $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function. Non-local interactions between cortical populations at positions $\mathbf{r}$ and $\mathbf{r}'$ are described by 
\begin{equation}
	\label{RateBasedInteractions} g\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)\, d\mathbf{r}'}, 
\end{equation}
where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial connectivity kernel and $\Omega$ is the spatial domain representing a cortical sheet or surface. The connectivity kernel is typically a ``Mexican hat'' or ``wizard hat'' function, which describes local excitation and surround inhibition. An example of the Mexican hat connectivity kernel used in this paper is shown in Figure~\ref{fig:Figure1}. To demonstrate flexibility in the estimation algorithm and to model larger cortical regions, a third component of the kernel is introduced describing weak longer-range excitation. The exact shape of this kernel is assumed to vary across subjects, and hence needs to be inferred from subject-specific data.

@@

\subsection{Derivation of Finite Dimensional State-Space Model}\label{Sect:ReducedModelDerivation}
In order to implement standard estimation techniques, we use a decomposition of the field using a set of Gaussian basis functions. Decomposition allows a continuous field to be represented by a finite-dimensional state vector. This facilitates application of standard nonlinear state estimation methods such as the unscented Kalman filter. The field decomposition is described by
\begin{equation}
	\label{DefFieldDecomp} v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t, 
\end{equation}
where $\mathbf{x}_t$ is the state vector that scales the field basis functions $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$. The superscript $\top$ denotes the transpose operator. The field basis functions are described by
\begin{equation}\label{eq:FieldBasisFunction}
	\phi\left(\mathbf{r}-\mathbf{r}'\right) =
\exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_{\phi}^2}\right)}, 
\end{equation}
where $\sigma_{\phi}$ is the basis function width parameter. An example of a one-dimensional field decomposition is given in Figure~\ref{fig:Figure2}.

@@

% Results and Discussion can be combined.
\section{Results}\label{ResultsSection} We now provide demonstrations of the estimation procedure using the new methods that are described above. Data was generated using the neural field model, described in Section~\ref{NeuralModelSection}, to demonstrate the state and parameter estimation framework. All parameters for the model are given in Table~\ref{tab:Parameters}, unless otherwise explicitly stated for particular experiments. Note that the parameters for the spatial and temporal aspects of the model can be considered arbitrary when demonstrating the estimation methodology. This is because the temporal aspects scale with the sampling rate and the length of the times series, and the spatial aspects scale with the size of the neural field and the spatial step size used for numerical simulation. When generating the simulated data, we assumed free boundary conditions for the neural field. An example of the simulated field at a single point in time can be seen in Figure~\ref{fig:Figure3}(A). Figure~\ref{fig:Figure3}(B) shows an example of data from the modelled sensors and Figure~\ref{fig:Figure3}(C) shows the power spectral density (PSD) of the simulated time-series. %The PSD shows the typical $1/f^2$ characteristics of intracranial EEG.

@@

\subsection{Spatial Frequency Analysis}
The spectral analysis was used to confirm that the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{fig:Figure4}(A) shows the spatial frequency of the neural field and Figure~\ref{fig:Figure4}(B) shows the spatial frequency of the reconstructed neural field from the basis function decomposition. The cutoff frequency ($-3$~dB point) of the neural field is $\boldsymbol{\nu}_c \approx 0.24$~Hz. From equation~\ref{eq:MinimumSensorDistance}, this yielded a maximum sensor separation of 2.08~mm. This confirms that the separation of $\Delta_{y} = 1.5$~mm was sufficient to mitigate problems associated with spatial aliasing of the dominant dynamics.

The spatial frequency of the observed field is shown in Figure~\ref{fig:Figure5}(A). The cutoff frequency is $\boldsymbol{\nu}_{cy} \approx 0.2$~Hz ($-3$~dB point). The observation cutoff frequency is lower than the field cutoff due to the low-pass filtering effect of the sensors. The spatial frequency response of the sensors can be seen in Figure~\ref{fig:Figure5}(B).

@@

Due to the limitations imposed by the computationally demanding estimation procedure we chose to use a $9\times9$ grid of Gaussian basis functions to decompose the field, which resulted in $81$ states.  Given the spatial domain, this yielded a basis function separation of $\Delta_{\phi}=2.5$~mm. The basis function width parameter was chosen such that a smooth representation of the field could be achieved with this arrangement, where $\sigma_{\phi}^2=2.5$. Given the basis function width, the decomposition is able to represent the field up to the spatial cutoff frequency of $\boldsymbol{\nu}_{c\phi} \approx 0.12$~Hz, as described by equation~\ref{eq:CutoffFromBasisFuncWidth}. The band-limiting effect is due to the spatial frequency response of the basis functions as seen in Figure~\ref{fig:Figure5}(C). The result of the band-limiting effect in the field reconstruction can be observed when comparing Figure~\ref{fig:Figure4}(A) to Figure~\ref{fig:Figure4}(B). Given the spatial cutoff frequency of the basis functions and the basis function separation the oversampling parameter is $\rho_{\phi}\approx 1.67$.

@@

Given this basis function separation, the basis function width parameter was set to $\sigma_{\phi}^2=2.5$. This allowed a smooth representation of the field by the basis function decomposition up to a spatial frequency $\boldsymbol{\nu}_{c\phi} \approx 0.12$~Hz, as described by equation~\ref{eq:CutoffFromBasisFuncWidth}. The band-limiting effect is due to the spatial frequency response of the basis functions as seen in Figure~\ref{fig:Figure5}(C). This band-limiting effect of the basis function decomposition can be observed when comparing Figure~\ref{fig:Figure4}(A) to Figure~\ref{fig:Figure4}(B).

@@

\subsubsection{Experiment I: Monte Carlo Simulation with Fixed Parameters}
A Monte Carlo approach was employed where 150 realisations of the data were generated, such that distributions of the parameters estimates could be formed. Each realisation consisted of 500~ms of data (sampled at $1$~kHz) and the estimation was applied to the final 400~ms, allowing the model's initial transients to die out. The initial state and parameters were unknown to the estimator. Ten iterations of the estimation algorithm were used to ensure that the parameter estimates converged. Figure~\ref{fig:Figure6} shows the rate of convergence, confirming that the algorithm converged to steady parameter values, with the changes in the parameter errors (between consecutive iterations) falling to less than $10^{-4}$ after 6 iterations.

Histograms of the parameter estimates obtained over the 150 realisations are shown in Figure~\ref{fig:Figure7}. In each of the subplots of the figure the true parameter value is shown by the solid line and the mean parameter estimate is shown by the dashed line. The actual parameter values, the mean parameter estimates, standard deviations, and biases for this parameter combination are given as the first entry in Table~\ref{tab:MexicanHatKernelResults}, where other entries in the table show results from Experiment II. The biases show the percentage differences between the actual values and the mean parameter estimates. For this parameter combination the mean estimates are in good agreement (within 1 standard deviation) with the actual values of the connectivity kernel parameters. However, the time constant parameter, $\xi$, is outside the estimate distribution (with bias of $2.56\%$), with a small variance in parameter estimates. We attribute this small bias to the loss of the high frequency components in the smoother estimated field with respect to the actual field. These results thus indicate that choosing to limit the frequency content of the reconstructed neural field has only minor effects on the accuracy of the estimated kernel while allowing for significant gains in the computational demand.

The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over iterations) of the root mean square error (MRMSE) over space. The MRMSE of the field estimates over time for 150 realisations is shown in Figure~\ref{fig:Figure9} with the $95~\%$ confidence interval. At each time point, the error is consistently around $0.5$~mV, which is approximately $10.73~\%$ of the field voltage range ([-2.20, 2.46]~mV). The error in the reconstructed field is again mostly due to the loss of high spatial frequency information from the basis function decomposition as illustrated in Figure~\ref{fig:Figure10}. The figure shows a snapshot at a single time instant of the true neural field, the estimated field, and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Figure~\ref{fig:Figure10}(B) that the estimated field contains the basic structure of the true field, shown in Figure~\ref{fig:Figure10}(A), but is missing the higher-frequency detail, which can be see in the error shown in Figure~\ref{fig:Figure10}(C).

@@

To study the robustness of the estimation algorithm, the performance was evaluated for various connectivity kernel parameters. The reconstructed connectivity kernels from the estimated parameters are shown in Figure~\ref{fig:Figure8}. For Figure~\ref{fig:Figure8}(A-C) the Mexican hat connectivity kernels were used. For each subset of parameters, the mean of the reconstructed kernel over 50 realisations is shown as a dotted line and the 95~\% confidence interval is indicated by the grey shaded region. In each case, the actual kernel was within the 95~\% confidence interval of the kernel estimate. The statistics for the kernel parameter estimates for each of the parameter sets are shown in Table~\ref{tab:MexicanHatKernelResults}.

For each of the Mexican hat connectivities, the support of the kernel was assumed to be known. However, in Figure~\ref{fig:Figure8}(D-F) the kernel support in the estimator differed from the actual data. In these cases the data sets were generated using a single positive (excitatory) Gaussian kernel, as seen by the solid line in each of the subplots of the figure. The estimation algorithm was constructed allowing for the Mexican hat connectivity of the previous examples. Each plot shows that the actual kernel lies within the confidence intervals of the reconstructed estimates for the majority of the kernel space, until the amplitude is relatively low. Table~\ref{tab:GaussianKernelResults} shows the statistics of the parameter estimates. Although the net reconstructed kernel estimates are in good accord with the actual kernels, the individual kernel basis function weights are not in agreement with the actual parameters. This result demonstrates that unless the widths of the Gaussians forming the kernel are precisely known, the interpretation of the results should be based on the net shape of the estimated connectivity kernel and not on the estimated parameter values individually.

@@

Figure~\ref{fig:Figure12} demonstrates the effect of the sigmoid parameters on the estimation accuracy. The RMSE of the reconstructed connectivity kernel was calculated for various parameters of the sigmoid by varying the slope, $\varsigma$, and the threshold, $v_0$. The other model parameters remained fixed and were set according to the Table~\ref{tab:Parameters}.

% The size of the squares in the figure show the magnitude of the log of the error. The squares are black if the log of the error is negative (the RMSE is $<1$) and the squares are white if the log of the error is positive (the RMSE is $>1$). The largest black squares indicate the best performance and the largest white squares indicate the worst performance for the parameter combinations (\dean{need some thing more quantitative}). 
From Figure~\ref{fig:Figure12} it can be seen that the estimation accuracy depends on the parameters of the sigmoid function. This is expected since by altering the parameters we are changing the nonlinearity and the saturation region of the sigmoid.

@@

\dean{PLEASE HELP TO FILL THIS SECTION IN. ANYTHING WILL DO.}
..\dean{This next bit is CRAP (please help):} When the threshold is negative or the slope is very gentle, all of the activity is being passed through the sigmoid function. However, as the slope is increased or the threshold is increased, more of the field is within the saturating region of the activation function. This is illustrated in Figure~\ref{fig:sigmoidfunctions}, where the sigmoids from the bottom left corner through to near top right corner of Figure~\ref{fig:sigmoidfunctions} are plotted. This appears to have a detrimental effect on the estimation accuracy.

\subsubsection{Experiment IV: Results for Varying the Sensor Kernel Width}
The accuracy of the estimation algorithm was tested for different observation kernels by varying the width parameter, $\sigma_m$. Figure~\ref{fig:Figure13} shows the RMSE of the connectivity kernel estimates for the various observation kernel widths. As the theory presented earlier in Section~\ref{SpectralAnalysisSection} suggests, the observation kernel acts as a spatial filter.

@@

When the observation kernel is too narrow, high frequency components in the neural field are sampled. These high frequency components result in aliasing errors in the estimated field thus affecting the kernel estimation accuracy. On the other hand, for the very wide kernels considered, important neural dynamics are not pick up by the estimated field due to the filtering out of the high frequency components by the sensors, resulting in poor accuracy in the kernel estimate. Thus, as indicated in Figure~\ref{fig:Figure13}, the best results are obtained when the sensor width is chosen to minimize both sources of errors.

@@

\subsection{Implications for Experimental Design}
As seen in Figures~\ref{fig:Figure7} and~\ref{fig:Figure10}, the results show good estimation accuracy for the connectivity kernels, where the true kernels lie within the 95~\% confidence intervals of the kernel estimates. The estimates of the connectivity kernels show relatively broad distributions over the different realisations, which can be expected considering the system's input is modelled as a random disturbance and the observations are corrupted by noise. To overcome the effect of noise, the use of a controlled input stimulus should be considered when applying the estimation algorithm to electrophysiological recordings, where given knowledge of the input and the appropriate model, it is expected that the variance of the parameter estimates would decrease. Even when using a known input, our results imply that we should only expect to learn distributions of parameters, rather than uncover a `true' parameter set. By designing an input signal carefully and using an evoked-potential paradigm where stimuli are repeated many times, recordings could be averaged over a sufficient number of trials (realisations), where each trial is a response to a stimulus, such that a distribution of the parameter estimates can be formed.

@@

\bibliographystyle{unsrt} 
\bibliography{BrainIDE}

@@

\section{Observation Kernel}
The Gaussian shape of the observation kernel was verified experimentally. This was achieved by moving a dipole across a recording electrode submerged in saline. A photograph through a microscope of the submerged electrodes can be seen in Figure~\ref{fig:SensorAndSource}. The measurement sensor was the most central electrode of a $5\times5$ Utah array (Blackrock Microsystems) which had electrode spacing of 400~$\mu$m. The current source was a twisted wire electrode (Plastics One Inc.) with tip-tip spacing of 60~$\mu$m and tip diameter of 200~$\mu$m. The twisted wire electrode was moved in the saline solution using a 3 dimensional manipulator (Sutter Instrument Company, MPC-100) in increments of 200~$\mu$m. The twisted wire electrode and the Utah array were co-registered visually using the microscope.
The stimulation waveform for the current source consisted of a 40~Hz sine wave with a duration of 1~s. The recording from the central electrode in the array was used to estimate the shape of the sensor kernel. To extract the amplitude of the signal, the data was first band-pass filtered with a centre frequency of 40~Hz. The envelope of the signal was found by computing the mean instantaneous amplitude over the 1~s stimulation duration by
\begin{equation}
	a = \frac{1}{N}\sum_n{\sqrt{(x(n)-\mathcal{H}x(n))^2}}
\end{equation}
where $\mathcal{H}$ denotes the Hilbert transform operator, $n$ indexes samples and $N$ is the total number of samples in the time series. Figure~\ref{fig:Ch3} shows the mean amplitude of the recorded envelope versus the distance between the twisted wire electrode and the recording electrode. A Gaussian fitted to the amplitudes is also shown. The figure shows that a Gaussian is indeed a good model for the sensor kernel in the homogeneous solution. 
\end{document}

>>>>

\subsection{Integro-Difference Equation Neural Field Model}
The combination of modelling techniques in this paper leads to a large amount of notation, so a reference of the symbols used is provided in Table~\ref{tab:Notation}. The model relates the average number of action potentials $g(\mathbf{r},t)$ arriving at time $t$ and position $\mathbf{r}$ to the local post-synaptic membrane voltage $v(\mathbf{r},t)$. The post-synaptic potentials generated at a neuronal population at location $\mathbf{r}$ by action potentials arriving from all other connected populations is described by 
\begin{equation}
	\label{SpikesToPotential} v\left( {\mathbf{r},t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right) \, dt'}. 
\end{equation}
The post-synaptic response kernel $h(t)$ is described by 
\begin{equation}
	\label{SynapticRespKernel} h(t) = \eta(t)\exp{\left(-\zeta t\right)}, 
\end{equation}
where $\zeta=\tau^{-1}$, $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function. Non-local interactions between cortical populations at positions $\mathbf{r}$ and $\mathbf{r}'$ are described by 
\begin{equation}
	\label{RateBasedInteractions} g\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)\, d\mathbf{r}'}, 
\end{equation}
where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial connectivity kernel and $\Omega$ is the spatial domain representing a cortical sheet or surface. The connectivity kernel is typically a ``Mexican hat'' or ``wizard hat'' function, which describes local excitation and surround inhibition. An example of the Mexican hat connectivity kernel used in this paper is shown in Fig.~\ref{fig:Figure1}. To demonstrate flexibility in the estimation algorithm and to model larger cortical regions, a third component of the kernel is introduced describing weak longer-range excitation. The exact shape of this kernel is assumed to vary across subjects, and hence needs to be inferred from subject-specific data.

@@

\subsection{Derivation of Finite Dimensional State-Space Model}\label{Sect:ReducedModelDerivation}
In order to implement standard estimation techniques, we use a decomposition of the field using a set of Gaussian basis functions. Decomposition allows a continuous field to be represented by a finite-dimensional state vector. This facilitates application of standard nonlinear state estimation methods such as the unscented Kalman filter. The field decomposition is described by
\begin{equation}
	\label{DefFieldDecomp} v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t, 
\end{equation}
where $\mathbf{x}_t$ is the state vector that scales the field basis functions $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$. The superscript $\top$ denotes the transpose operator. The field basis functions are described by
\begin{equation}\label{eq:FieldBasisFunction}
	\phi\left(\mathbf{r}-\mathbf{r}'\right) =
\exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_{\phi}^2}\right)}, 
\end{equation}
where $\sigma_{\phi}$ is the basis function width parameter. An example of a one-dimensional field decomposition is given in Fig.~\ref{fig:Figure2}.

@@

% Results and Discussion can be combined.
\section{Results}\label{ResultsSection} We now provide demonstrations of the estimation procedure using the new methods that are described above. Data was generated using the neural field model, described in Section~\ref{NeuralModelSection}, to demonstrate the state and parameter estimation framework. All parameters for the model are given in Table~\ref{tab:Parameters}, unless otherwise explicitly stated for particular experiments. Note that the parameters for the spatial and temporal aspects of the model can be considered arbitrary when demonstrating the estimation methodology. This is because the temporal aspects scale with the sampling rate and the length of the times series, and the spatial aspects scale with the size of the neural field and the spatial step size used for numerical simulation. When generating the simulated data, we assumed free boundary conditions for the neural field. An example of the simulated field at a single point in time can be seen in Fig.~\ref{fig:Figure3}(A). Fig.~\ref{fig:Figure3}(B) shows an example of data from the modelled sensors and Fig.~\ref{fig:Figure3}(C) shows the power spectral density (PSD) of the simulated time-series. %The PSD shows the typical $1/f^2$ characteristics of intracranial EEG.

@@

\subsection{Spatial Frequency Analysis}
The spectral analysis was used to confirm that the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Fig.~\ref{fig:Figure4}(A) shows the spatial frequency of the neural field and Fig.~\ref{fig:Figure4}(B) shows the spatial frequency of the reconstructed neural field from the basis function decomposition. The cutoff frequency ($-3$~dB point) of the neural field is $\boldsymbol{\nu}_c \approx 0.24$~Hz. From equation~\ref{eq:MinimumSensorDistance}, this yielded a maximum sensor separation of 2.08~mm. This confirms that the separation of $\Delta_{y} = 1.5$~mm was sufficient to mitigate problems associated with spatial aliasing of the dominant dynamics.

The spatial frequency of the observed field is shown in Fig.~\ref{fig:Figure5}(A). The cutoff frequency is $\boldsymbol{\nu}_{cy} \approx 0.2$~Hz ($-3$~dB point). The observation cutoff frequency is lower than the field cutoff due to the low-pass filtering effect of the sensors. The spatial frequency response of the sensors can be seen in Fig.~\ref{fig:Figure5}(B).

@@

Due to the limitations imposed by the computationally demanding estimation procedure we chose to use a $9\times9$ grid of Gaussian basis functions to decompose the field, which resulted in $81$ states.  Given the spatial domain, this yielded a basis function separation of $\Delta_{\phi}=2.5$~mm. The basis function width parameter was chosen such that a smooth representation of the field could be achieved with this arrangement, where $\sigma_{\phi}^2=2.5$. Given the basis function width, the decomposition is able to represent the field up to the spatial cutoff frequency of $\boldsymbol{\nu}_{c\phi} \approx 0.12$~Hz, as described by equation~\ref{eq:CutoffFromBasisFuncWidth}. The band-limiting effect is due to the spatial frequency response of the basis functions as seen in Fig.~\ref{fig:Figure5}(C). The result of the band-limiting effect in the field reconstruction can be observed when comparing Fig.~\ref{fig:Figure4}(A) to Fig.~\ref{fig:Figure4}(B). Given the spatial cutoff frequency of the basis functions and the basis function separation the oversampling parameter is $\rho_{\phi}\approx 1.67$.

@@

Given this basis function separation, the basis function width parameter was set to $\sigma_{\phi}^2=2.5$. This allowed a smooth representation of the field by the basis function decomposition up to a spatial frequency $\boldsymbol{\nu}_{c\phi} \approx 0.12$~Hz, as described by equation~\ref{eq:CutoffFromBasisFuncWidth}. The band-limiting effect is due to the spatial frequency response of the basis functions as seen in Fig.~\ref{fig:Figure5}(C). This band-limiting effect of the basis function decomposition can be observed when comparing Fig.~\ref{fig:Figure4}(A) to Fig.~\ref{fig:Figure4}(B).

@@

\subsubsection{Experiment I: Monte Carlo Simulation with Fixed Parameters}
A Monte Carlo approach was employed where 150 realisations of the data were generated, such that distributions of the parameters estimates could be formed. Each realisation consisted of 500~ms of data (sampled at $1$~kHz) and the estimation was applied to the final 400~ms, allowing the model's initial transients to die out. The initial state and parameters were unknown to the estimator. Ten iterations of the estimation algorithm were used to ensure that the parameter estimates converged. Fig.~\ref{fig:Figure6} shows the rate of convergence, confirming that the algorithm converged to steady parameter values, with the changes in the parameter errors (between consecutive iterations) falling to less than $10^{-4}$ after 6 iterations.

Histograms of the parameter estimates obtained over the 150 realisations are shown in Fig.~\ref{fig:Figure7}. In each of the subplots of the figure the true parameter value is shown by the solid line and the mean parameter estimate is shown by the dashed line. The actual parameter values, the mean parameter estimates, standard deviations, and biases for this parameter combination are given as the first entry in Table~\ref{tab:MexicanHatKernelResults}, where other entries in the table show results from Experiment II. The biases show the percentage differences between the actual values and the mean parameter estimates. For this parameter combination the mean estimates are in good agreement (within 1 standard deviation) with the actual values of the connectivity kernel parameters. However, the time constant parameter, $\xi$, is outside the estimate distribution (with bias of $2.56\%$), with a small variance in parameter estimates. We attribute this small bias to the loss of the high frequency components in the smoother estimated field with respect to the actual field. These results thus indicate that choosing to limit the frequency content of the reconstructed neural field has only minor effects on the accuracy of the estimated kernel while allowing for significant gains in the computational demand.

The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over iterations) of the root mean square error (MRMSE) over space. The MRMSE of the field estimates over time for 150 realisations is shown in Fig.~\ref{fig:Figure9} with the $95~\%$ confidence interval. At each time point, the error is consistently around $0.5$~mV, which is approximately $10.73~\%$ of the field voltage range ([-2.20, 2.46]~mV). The error in the reconstructed field is again mostly due to the loss of high spatial frequency information from the basis function decomposition as illustrated in Fig.~\ref{fig:Figure10}. The figure shows a snapshot at a single time instant of the true neural field, the estimated field, and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Fig.~\ref{fig:Figure10}(B) that the estimated field contains the basic structure of the true field, shown in Fig.~\ref{fig:Figure10}(A), but is missing the higher-frequency detail, which can be see in the error shown in Fig.~\ref{fig:Figure10}(C).

@@

To study the robustness of the estimation algorithm, the performance was evaluated for various connectivity kernel parameters. The reconstructed connectivity kernels from the estimated parameters are shown in Fig.~\ref{fig:Figure8}. For Fig.~\ref{fig:Figure8}(A-C) the Mexican hat connectivity kernels were used. For each subset of parameters, the mean of the reconstructed kernel over 50 realisations is shown as a dotted line and the 95~\% confidence interval is indicated by the grey shaded region. In each case, the actual kernel was within the 95~\% confidence interval of the kernel estimate. The statistics for the kernel parameter estimates for each of the parameter sets are shown in Table~\ref{tab:MexicanHatKernelResults}.

For each of the Mexican hat connectivities, the support of the kernel was assumed to be known. However, in Fig.~\ref{fig:Figure8}(D-F) the kernel support in the estimator differed from the actual data. In these cases the data sets were generated using a single positive (excitatory) Gaussian kernel, as seen by the solid line in each of the subplots of the figure. The estimation algorithm was constructed allowing for the Mexican hat connectivity of the previous examples. Each plot shows that the actual kernel lies within the confidence intervals of the reconstructed estimates for the majority of the kernel space, until the amplitude is relatively low. Table~\ref{tab:GaussianKernelResults} shows the statistics of the parameter estimates. Although the net reconstructed kernel estimates are in good accord with the actual kernels, the individual kernel basis function weights are not in agreement with the actual parameters. This result demonstrates that unless the widths of the Gaussians forming the kernel are precisely known, the interpretation of the results should be based on the net shape of the estimated connectivity kernel and not on the estimated parameter values individually.

@@

Fig.~\ref{fig:Figure12} demonstrates the effect of the sigmoid parameters on the estimation accuracy. The RMSE of the reconstructed connectivity kernel was calculated for various parameters of the sigmoid by varying the slope, $\varsigma$, and the threshold, $v_0$. The other model parameters remained fixed and were set according to the Table~\ref{tab:Parameters}.

% The size of the squares in the figure show the magnitude of the log of the error. The squares are black if the log of the error is negative (the RMSE is $<1$) and the squares are white if the log of the error is positive (the RMSE is $>1$). The largest black squares indicate the best performance and the largest white squares indicate the worst performance for the parameter combinations (\dean{need some thing more quantitative}). 
From Fig.~\ref{fig:Figure12} it can be seen that the estimation accuracy depends on the parameters of the sigmoid function. This is expected since by altering the parameters we are changing the nonlinearity and the saturation region of the sigmoid.

@@

\dean{PLEASE HELP TO FILL THIS SECTION IN. ANYTHING WILL DO.}
..\dean{This next bit is CRAP (please help):} When the threshold is negative or the slope is very gentle, all of the activity is being passed through the sigmoid function. However, as the slope is increased or the threshold is increased, more of the field is within the saturating region of the activation function. This is illustrated in Fig.~\ref{fig:sigmoidfunctions}, where the sigmoids from the bottom left corner through to near top right corner of Fig.~\ref{fig:sigmoidfunctions} are plotted. This appears to have a detrimental effect on the estimation accuracy.

\subsubsection{Experiment IV: Results for Varying the Sensor Kernel Width}
The accuracy of the estimation algorithm was tested for different observation kernels by varying the width parameter, $\sigma_m$. Fig.~\ref{fig:Figure13} shows the RMSE of the connectivity kernel estimates for the various observation kernel widths. As the theory presented earlier in Section~\ref{SpectralAnalysisSection} suggests, the observation kernel acts as a spatial filter.

@@

When the observation kernel is too narrow, high frequency components in the neural field are sampled. These high frequency components result in aliasing errors in the estimated field thus affecting the kernel estimation accuracy. On the other hand, for the very wide kernels considered, important neural dynamics are not pick up by the estimated field due to the filtering out of the high frequency components by the sensors, resulting in poor accuracy in the kernel estimate. Thus, as indicated in Fig.~\ref{fig:Figure13}, the best results are obtained when the sensor width is chosen to minimize both sources of errors.

@@

\subsection{Implications for Experimental Design}
As seen in Fig.~\ref{fig:Figure7} and Fig.~\ref{fig:Figure10}, the results show good estimation accuracy for the connectivity kernels, where the true kernels lie within the 95~\% confidence intervals of the kernel estimates. The estimates of the connectivity kernels show relatively broad distributions over the different realisations, which can be expected considering the system's input is modelled as a random disturbance and the observations are corrupted by noise. To overcome the effect of noise, the use of a controlled input stimulus should be considered when applying the estimation algorithm to electrophysiological recordings, where given knowledge of the input and the appropriate model, it is expected that the variance of the parameter estimates would decrease. Even when using a known input, our results imply that we should only expect to learn distributions of parameters, rather than uncover a `true' parameter set. By designing an input signal carefully and using an evoked-potential paradigm where stimuli are repeated many times, recordings could be averaged over a sufficient number of trials (realisations), where each trial is a response to a stimulus, such that a distribution of the parameter estimates can be formed.

@@

\bibliographystyle{unsrt} 
\bibliography{BrainIDE}

@@

\section{Observation Kernel}
The Gaussian shape of the observation kernel was verified experimentally. This was achieved by moving a dipole across a recording electrode submerged in saline. A photograph through a microscope of the submerged electrodes can be seen in Fig.~\ref{fig:SensorAndSource}. The measurement sensor was the most central electrode of a $5\times5$ Utah array (Blackrock Microsystems) which had electrode spacing of 400~$\mu$m. The current source was a twisted wire electrode (Plastics One Inc.) with tip-tip spacing of 60~$\mu$m and tip diameter of 200~$\mu$m. The twisted wire electrode was moved in the saline solution using a 3 dimensional manipulator (Sutter Instrument Company, MPC-100) in increments of 200~$\mu$m. The twisted wire electrode and the Utah array were co-registered visually using the microscope.
The stimulation waveform for the current source consisted of a 40~Hz sine wave with a duration of 1~s. The recording from the central electrode in the array was used to estimate the shape of the sensor kernel. To extract the amplitude of the signal, the data was first band-pass filtered with a centre frequency of 40~Hz. The envelope of the signal was found by computing the mean instantaneous amplitude over the 1~s stimulation duration by
\begin{equation}
	a = \frac{1}{N}\sum_n{\sqrt{(x(n)-\mathcal{H}x(n))^2}}
\end{equation}
where $\mathcal{H}$ denotes the Hilbert transform operator, $n$ indexes samples and $N$ is the total number of samples in the time series. Fig.~\ref{fig:Ch3} shows the mean amplitude of the recorded envelope versus the distance between the twisted wire electrode and the recording electrode. A Gaussian fitted to the amplitudes is also shown. The figure shows that a Gaussian is indeed a good model for the sensor kernel in the homogeneous solution. 
\end{document}
